Global Distribution of Unintended Pregnancies and Contraceptive Usage

### 2024-04-15 code/eda-code/eda.qmd
Outcome Description
Updated our outcome variable description in eda and manuscript files to better explain our results.

Plot 3 uses `pct_upreg_sarc` and `sarc_standard` to examine the relationship with short-acting reversible methods. 
```{r}

# explore percentage of unintended pregnancies compared to standardized count of women using short acting reversible methods
plot_3 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Percentage of Unintended Pregnancies") + ylab("Standarized Total Number of Women")
plot(plot_3)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png")
ggsave(filename = figure_file, plot=plot_3)

@@

Plot 3: The same positive trend is shown between the percentage of unintended pregnancies and number of women using SARMs, but there is considerably more variation in this relationship.

>>>>

Plot 3 uses `pct_upreg_sarc` and `sarc_standard` to examine the relationship with short-acting reversible methods. 
```{r}
# explore percentage of unintended pregnancies compared to standardized count of women using short acting reversible methods
plot_3 <- mydata %>% ggplot(aes(x=sarc_standard, y=pct_upreg_sarc)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Standarized Total Number of Women") + ylab("Percentage of Unintended Pregnancies")
plot(plot_3)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png")
ggsave(filename = figure_file, plot=plot_3)

@@

Plot 3: The same positive trend is shown between the percentage of unintended pregnancies and number of women using SARCs, but there is considerably more variation in this relationship. The outcome of interest should increase as the number of women using short acting-reversible contraception increases because more women are at risk of experiencing an unintended pregnancy due to this contraception if they are using this type of contraception in the first place. This percentage does not reflect the effectiveness of short-acting reversible contraception, but it does classify the reason for an unintended pregnancy. As explained in the supplementary material, the dataset includes the percentage of unintended pregnancies among long-acting reversible contraception and sterilization users, traditional birth control users, and users who do not use a method of birth control. It makes sense that our outcome would increase as more women put themselves at risk of experiencing the outcome through becoming classified as short-acting reversible contraception users. This association is a little challenging to contextualize since we could not analyze all 4 classifications for the percentage of unintended pregnancies.

### 2024-04-15 products/manuscript/Manuscript.qmd
Outcome Description
Updated our outcome variable description in eda and manuscript files to better explain our results.

The next set of variables we explored were the total counts of women using each birth control method. There were two large outliers that prevented the graphs from being useful. After exploring the outliers, we standardized using the population size of reproductive aged women (15-49 years). There is a positive association between the percentage of unintended pregnancies among short acting reversible method users and the total number of women using that method of birth control, which is to be expected. Our outcome variable is the number of unintended pregnancies among women who are using short-acting reversible contraception, so the number of unintended pregnancies that can be classified in this outcome is dependent on the number of women using this type of contraception. This percentage should increase as the number of women using short acting-reversible contraception increases because more women are at risk of experiencing an unintended pregnancy due to this contraception if they are using this type of contraception in the first place. This outcome variable does not speak to the effectiveness of short-acting reversible contraception, so it makes sense that our outcome would increase as more women put themselves at risk of experiencing the outcome through becoming classified as short-acting reversible contraception users. This association is displayed in @exploratory-result2.
```{r}
#| label: exploratory-result2
#| fig-cap: "Figure 1: Graph Depciting Percent Unintended Pregnancy by Standarized Count of Total Women"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png"))
```

>>>>

The next set of variables we explored were the total counts of women using each birth control method. There were two large outliers that prevented the graphs from being useful. After exploring the outliers, we standardized using the population size of reproductive aged women (15-49 years). There is a positive association between the percentage of unintended pregnancies among short acting reversible method users and the total number of women using that method of birth control, which is to be expected. Our outcome variable is the number of unintended pregnancies among women who are using short-acting reversible contraception, so the number of unintended pregnancies that can be classified in this outcome is dependent on the number of women using this type of contraception. This percentage should increase as the number of women using short acting-reversible contraception increases because more women are at risk of experiencing an unintended pregnancy due to this contraception if they are using this type of contraception in the first place. This outcome variable does not speak to the effectiveness of short-acting reversible contraception, so it makes sense that our outcome would increase as more women put themselves at risk of experiencing the outcome through becoming classified as short-acting reversible contraception users. As explained in the supplementary material, the dataset includes the percentage of unintended pregnancies among long-acting reversible contraception and sterilization users, traditional birth control users, and users who do not use a method of birth control. Each of these percentages are positively associated with number of people in its respective classification group that it represents. The association for the contraception that we chose to focus on for this analysis, short acting reversible contraception, is displayed in @exploratory-result2.
```{r}
#| label: exploratory-result2
#| fig-cap: "Figure 1: Graph Depciting Percent Unintended Pregnancy by Standarized Count of Total Women"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png"))
```
